---
title: "Genitive Alternation Learners"
output:
  html_document:
    keep_md: true
date: "November 28, 2017"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.kable.NA = '')
```

## Introduction

We will explore **<genitivealternation_learners.csv>** to see if we can predict which genitive marker learners use by exploring a few predictors and their interaction.  We will explore using a classification tree model.

### Imports & Initial Data Check

```{r, message=FALSE}
#imports
library("tidyverse"); library("car"); library("effects"); library("tree"); library("knitr")

x <- read.delim("genitivealternation_learners.csv")

#Remove variables we will not be considering
x$LENGTHNP1 <- NULL
x$LENGTHNP2 <- NULL
attach(x)
```

After loading the appropriate libraries & data we look at the first and last five rows.





```{r}

kable(head(x), caption = "First Five Rows")
kable(head(x), caption = "Last Five Rows")
kable(summary(x), caption = "Summary")
```

Immediately looking at the summary, we can see an imbalance between the observations. Some categorical predictors are poorly represented in the data such as possessor animacy (POSSORANIM), where the category *animate* is only found twice in the data set.


```{r}
ggplot(data = x, aes(x = MATCH, fill = MATCH)) +
geom_bar()
prop.table(table(x$MATCH))

```
We can see that observations of the genitive "of" outnumber 's' practically 9 to 1.  With this being said, a basic model of just picking 'of' for every example would give our model an accuracy of 88% from the beginning.  This implies that 88% is not an acceptable model to really explain what is happening.


```{r}

prop.table(table(x$MATCH))
```




```{r}
tree.1 <- tree(MATCH ~ SEMCLASS + POSSORANIM + POSSEDANIM + POSSORSPEC + POSSEDSPEC + POSSORBRANCH + POSSEDBRANCH + POSSORCOMPL + POSSEDCOMPL)


summary(tree.1)
```
It seems that all predictors are relevant except POSSEDANIM, POSSORCOMPL, and POSSEDCOMPL.

Next we look at the effects numerically.

```{r}
tree.1

```

We plot the tree.

```{r}
plot(tree.1)
text(tree.1, pretty = 4, all = T)

```

# Model Validation

To test help evaluate our tree, we decided to preform a simple cross validation.

```{r}
set.seed(42)

sampler <- sample(rep(c("train", "test"), c(891,99)))

#Generate a model, that is fitted to only the testing data
validation <- tree(formula(tree.1), data = x[sampler == "train",])

#Have the model predict labels of the "train" set
validation.pred <- predict(validation, newdata = x[sampler == "test",], type = "class")

#Accuracy Score
sum(validation.pred == MATCH[sampler == "test"]) / length(validation.pred)



```



## Cross Validation & Pruning

```{r}
#Generate a pruned model
pruning <- cv.tree(tree.1, FUN = prune.misclass)
pruning

```


```{r}
plot(pruning$size, pruning$dev, type="b", 
     xlab = "Number of Nodes", 
     ylab = "Deviation")
grid()

```

The deviances are lowest for 5, 6, and 10 nodes, we pick 5 (because 6 is higher than 5 and 10 is the original number of nodes). We then prune our  `tree.1` model and plot it.

```{r}

tree.2 <- prune.misclass(tree.1, best = 5)

plot(tree.2)
text(tree.2, pretty = 0, all = T)
```


### Pruned Model Validation

We run the same validation tests as before using the same test/train set.

```{r}

#Generate a model, that is fitted to only the testing data
validation.2 <- tree(formula(tree.2), data = x[sampler == "train",])

#Predict labels of the "train" set
validation.pred.2 <- predict(validation.2, newdata = x[sampler == "test",], type = "class")

#Accuracy Score
sum(validation.pred.2 == MATCH[sampler == "test"]) / length(validation.pred.2)


kable(table(MATCH[sampler == "test"], validation.pred.2), caption = "Confusion Matrix")

```

```{r}

#Compute Values
accuracy <- (82 + 8) / length(validation.pred.2)
precision <- 8/ (8 + 6)
recall <- 8 / (8 + 3)
f.score <- 2 * ((precision * recall))/ (precision + recall)

#Set-up Data-Frame
Metric <- c("Accuracy:", "Precision:", "Recall:", "F-Score:")
Score <- c(accuracy, precision, recall, f.score)
df <- data.frame(Metric,Score)

#Show Table
kable(df, caption = "Validation for Pruned Tree")

```

# Conclusion

The match will be “of" unless...

1. the specificity of the possessed is nonspecific and the branching of the possessor is not post-modified, pre-modified, or prepostmodified, and the specificity of the possessor is nonspecific.

OR 

2. the specificity of the possessed is not nonspecific and the branching of the possessed is not none or premodified, then the match will be " 's ".


 





